\begin{thebibliography}{12}
\bibitem{AntRDS}
Bergua A., Skrandiesb W., 2000,
\textit{An early antecedent to modern random dot stereograms - 'The Secret Stereoscopic Writing' of Ram\'on y Cajal}.
Int. J. of Psychophysiology 36, 69-72\\
\url{http://dx.doi.org/10.1016/S0167-8760(99)00111-7}

\bibitem{BellRDS}
Julesz B., 1960.
\textit{Binocular depth perception of computer-generated patterns}.
Bell Syst. Tech. J. 39, 1125-1162.\\
\url{http://doi.apa.org/?uid=1961-01385-001}

\bibitem{GenRDS}
Gonzalez F., Krause F., 1994.
\textit{Generation of dynamic random-element stereograms in real time with a system based on a personal computer}.
Med. \& Biol. Eng. \& Comput., 1994, 32, 373-376.\\
\url{http://dx.doi.org/10.1007/BF02524687}

\bibitem{DepthCues}
Cutting J., Vishton P., 1995.
\textit{Perceiving layout and knowing distances: The integration, relative potency, and contextual use of different information about depth}.
In W. Epstein \& S. Rogers (Eds.), Handbook of perception and cognition, Vol. 5, 69-117. San Diego, CA: Academic Press.\\
\url{http://dx.doi.org/10.1016/B978-012240530-3/50005-5}

\bibitem{MeshPop}
Bischoff Stephan, Kobbelt Leif, 2004.
\textit{Teaching meshes, subdivision and multiresolution techniques}.
Computer-Aided DesignVolume 36, Issue 14, , CAD Education, December 2004, Pages 1483-1500.\\
\url{http://dx.doi.org/10.1016/j.cad.2003.11.007}
%Bischoff S., Kobbelt L., 2004.
%\textit{Teaching meshes, subdivision and multiresolution techniques}.
%Computer-Aided DesignVolume 36, Issue 14, , CAD Education, December 2004, Pages 1483-1500.\\
%\url{http://dx.doi.org/10.1016/j.cad.2003.11.007}

\bibitem{Mayan}
R\"othlin Gerhard, 2007.
\textit{Semester Thesis: Mayan, A novel algorithm to create color preserving anaglyph images}.
Semester Thesis, 2007.

\end{thebibliography}

% Might only work in KOMA script (scrarticle)
\renewcommand*\refname{Links}

\begin{thebibliography}{12}
\bibitem[a]{proj} \url{http://en.wikipedia.org/wiki/Graphical_projection}
\bibitem[b]{parallel} \url{http://en.wikipedia.org/wiki/Parallel_projection}
\bibitem[c]{perspective} \url{http://en.wikipedia.org/wiki/Perspective_projection}
\bibitem[d]{painters} \url{http://en.wikipedia.org/wiki/Painter's_algorithm}
\bibitem[e]{zbuffer} \url{http://en.wikipedia.org/wiki/Z-buffering}
\bibitem[f]{fog} \url{http://www.opengl.org/documentation/specs/version1.1/glspec1.1/node90.html}
\bibitem[g]{accommodation} \url{http://en.wikipedia.org/wiki/Accommodation_(eye)}
\bibitem[h]{triangulation} \url{http://en.wikipedia.org/wiki/Triangulation}
\bibitem[i]{scenegraph} \url{http://www.realityprime.com/articles/scenegraphs-past-present-and-future}
\bibitem[j]{turing} \url{http://en.wikipedia.org/wiki/Turing_completeness}
\bibitem[k]{procedural} \url{http://en.wikipedia.org/wiki/Procedural_programming}
\bibitem[l]{dataflow} \url{http://en.wikipedia.org/wiki/Dataflow_programming}
\bibitem[m]{fsm} \url{http://en.wikipedia.org/wiki/Finite_state_machine}
\bibitem[n]{lua} \url{http://www.lua.org/manual/5.1/manual.html}
\bibitem[o]{boost} \url{http://www.boost.org/}
\bibitem[p]{git} \url{http://git.or.cz/}
\bibitem[q]{qt} \url{http://trolltech.com/downloads/opensource}
\bibitem[r]{lb} \url{http://luabridge.sourceforge.net/}
\bibitem[s]{hig} \url{http://developer.apple.com/documentation/UserExperience/Conceptual/AppleHIGuidelines/}
\bibitem[t]{deployment} \url{http://doc.trolltech.com/4.4/deployment.html}
\bibitem[u]{opengl} \url{http://en.wikipedia.org/wiki/OpenGL}
\bibitem[v]{libobj} \url{http://www.cs.kuleuven.ac.be/~ares/libobj/index.html}
\bibitem[w]{superlab} \url{http://superlab.com/}
\bibitem[z]{project} \url{http://the-color-black.net/examinationroom/}
\end{thebibliography}
